\sysfull

### 2018-08-02 paper/asplos19/paper/intro.tex
preparing for asplos19

Advances in processor efficiency along with the development of
energy-harvesting systems have created a new category of devices that require
neither a battery nor a tethered power
supply~\citep{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}. These
devices operate using ambient energy, such as radio frequency
transmissions~\citep{rf_powered_computing_gollakota_2014},
light~\citep{margolies_infocom_2016,margolies_tosn_2016}, and
vibration~\citep{gorlatova_sigmetrics_2014}. Incorporating compute, storage,
sensing, and communication hardware~\citep{wisp5,moo,capybara}, such devices are a
promising technology for use in the Internet of Things~\citep{ku_cst_2016},
in-body~\citep{nadeau_naturebio_2017} and
on-body~\citep{bandodkar_electroanalysis_2015} medical systems, and
energy-harvesting nano-satellites~\citep{kicksat,capybara}.

Energy-harvesting devices create unique challenges because they operate {\em
intermittently} when energy is
available~\citep{hicks_isca_2017,lucia_snapl_2017}. An energy-harvesting device
buffers energy in a small storage capacitor~\citep{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., nW to $\mu$W) compared to a platform's operating
power level (hundreds of $\mu$W to mW). A device operates briefly until it depletes its buffered energy, after which, the device shuts
down and recharges to operate again later---corresponds to {\em intermittent execution model}~\citep{dino,lucia_snapl_2017} composed of
operation-power failure-restart cycles. As an example, recharge times may be
tens of seconds in radio frequency powered medical device~\cite[Fig.
3c]{nadeau_naturebio_2017}. The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor~\cite{capybara} and some devices fail and restart operating $\approx$10 to
$\approx$100 times per second~\citep{tan_infocom_2016,mementos,nvp}.

\textbf{Data Consistency in Intermittent Computing.}  
%Software in an energy-harvesting system operates according to the {\em intermittent execution model}~\citep{dino,lucia_snapl_2017}, which corresponds to the operation-failure-restart cycles of an energy-harvesting device. In an intermittent software execution, an {\em operating period} proceeds for anarbitrary duration (dictated by energy availability) before being interruptedby a {\em failure period}. In a failure period, 
Upon power failures, a device loses the volatile
state in its registers, stack, SRAM, and retains the state of any non-volatile
memory, such as FRAM or Flash. While capturing periodic
checkpoints~\citep{mementos,quickrecall} and sleep
scheduling~\citep{dewdrop,hibernus,hibernusplusplus} help preserve execution
progress, failures can leave non-volatile state incorrectly, partially updated,
and may leave checkpointed volatile state inconsistent with non-volatile state.
These inconsistencies cause intermittent execution to deviate from
continuously-powered behavior, often leading to an unrecoverable
application failure~\citep{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for
intermittently-powered devices: (i) \emph{software-based programming and
execution models}~\citep{dino,ratchet,chain,alpaca} and (ii)
\emph{hardware-based computer architecture
support}~\citep{hicks_isca_2017,idetic,nvp}. Complex architectural changes are
expensive to design, verify, and manufacture. New architectures are also
inapplicable to existing systems~\citep{hicks_isca_2017,nvp}. Software
approaches are simpler and applicable to existing devices today, but prior
software approaches have a number of limitations that are the focus of the work
in this paper. A key limitation of prior software approaches that we address in this paper is the {\em inflexibility} of a program statically decomposed into tasks.

\begin{wrapfigure}{t!}{0.45\textwidth}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/intro-figure.pdf}
    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure). On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
    \label{fig:coalesce}
\end{wrapfigure}

\textbf{Task Decomposition of Intermittent Programs.} {\em
Task-based} programming and execution models require a
programmer~\citep{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to
statically decompose a program into a collection of {\em tasks}.  A task can
include arbitrary computation that should be executed despite arbitrarily-timed power failures.
%and these existing systems guarantee that each
%tasks executes {\em atomically}, despite arbitrarily-timed power failures.  
%
The programmer explicitly expresses task-to-task control flow, or the compiler
converts existing program control-flow into task-to-task control flow to sequence 
the tasks that it defines.
%
Figure~\ref{fig:coalesce} illustrates how a program's tasks execute and
shows how tasks can impose a run-time overhead. At each task transition, the system incurs an overhead to track and atomically \emph{commit} modifications to the non-volatile memory, to maintain consistency of program state~\citep{chain,alpaca}. The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.

@@

%A key challenge is that the length of a software
%task's execution is \emph{limited by the fixed total amount of energy} that a
%device can buffer in hardware. A task's code is static, but the duration of its
%execution may be input-dependent and is difficult to predict. To illustrate
%this, we refer to Figure~\ref{fig:1} showing the execution time of two
%applications (the same application X, but divided into X and Y tasks as
%in~\citep{chain}) running on RFID antenna-powered Computational
%RFID~\citep{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna
%distances (near---stable energy supply/far---high energy intermittency). At
%short distance: execution will takes too long caused by runtime cost of
%marshalling excessively small tasks. At far distance: program \emph{might never
%execute} if task execution consumes more energy than the system can buffer.
%This calls for at-runtime adaptive task division of any transiently-powered
%application---namely \emph{task virtualization}.

@@

%variables in non-volatile memory at \emph{run-time} that can
%break the atomicity of the \emph{virtual task}. Consider the
%example depicted in Fig.~\ref{fig:virtualization}: Tasks 1,
%2 and 3 being executed consecutively. All tasks in this
%example are atomic since they do not have WAR dependency on
%the persistent variables they are accessing, e.g. \emph{x}
%is only read and \emph{y} is only written within Task 1.
%Now, suppose three tasks have been virtualized into a single
%one at run-time, namely Task 4: since \emph{x} is now first
%\emph{read} and then \emph{written}, a WAR dependency on
%\emph{x} is introduced dynamically at
%run-time---unfortunately Task 4 is \emph{no more atomic}
%since its re-execution will not always produce the same
%results.
\textbf{Challenges.} While a compiler can use static data privatization and commit instrumentation
(i.e., redo-logging) to eliminate statically identifiable, inter-task data
dependences~\citep{alpaca}, a \emph{dynamic} dependence between two coalesced
tasks requires dynamic privatization and commit actions. 
%
To ensure consistency and respect inter-task dependences, \sys uses a novel
approach \emph{power failure-immune memory virtualization} that buffers non-volatile variable
updates in volatile memory during a coalesced task execution, before committing
them to non-volatile memory at the dynamic task boundary.
%
%Therefore, we require a \emph{new execution
%model} that keeps each virtual task atomic by committing the
%modified persistent variables at the boundary defined
%dynamically at run-time---keeping the non-volatile memory
%unmodified upon a power-interrupt and preserving its
%consistency.
A static task decomposition model assumes that each
single task can execute to completion. If the hardware energy buffer provides
inadequate energy to execute each single task to completion, a program will not
terminate~\cite{cleancut_2018}. To avoid non-termination under adversarial
energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
Partial execution avoids non-termination by committing the intermediate state of a long-running
task that has repeatedly failed and restarted. Partial execution violates
task atomicity, but preserves forward progress; if a programmer knows that
task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
%
% !!!! % It should be emphasized that unlike a voltage-threshold-based protection approaches that may suffer from a non-termination problem if the checkpoint has to be placed within the interrupt-disabled region, a timer-based approach adapts its partial commit location on repeated failures. 
%
%Task coalescing removes the burden from the programmer, because it accepts any
%task decomposition and improves it dynamically. To further reduce the
%programmer effort, we propose a compiler pass for \emph{automatic
%decomposition} of programs into (small) \emph{atomic} tasks. The compiler
%identifies non-volatile variables shared across tasks as tasks are created, and
%instruments reads and writes of those variables, using memory virtualization,
%to keep the data consistent in the presence of power loss.
%
%\textcolor{red}{Despite being limited to a subset of the C language, the
%automatic task decomposition allowed us to port several	applications to an
%intermittent platform with a moderate effort.}
%

>>>>

Advances in processor efficiency along with the development of
energy-harvesting systems have created a new category of devices that require
neither a battery nor a tethered power
supply~\cite{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}. These
devices operate using ambient energy, such as radio frequency
transmissions~\cite{rf_powered_computing_gollakota_2014},
light~\cite{margolies_infocom_2016,margolies_tosn_2016}, and
vibration~\cite{gorlatova_sigmetrics_2014}. Incorporating compute, storage,
sensing, and communication hardware~\cite{wisp5,moo,capybara}, such devices are a
promising technology for use in the Internet of Things~\cite{ku_cst_2016},
in-body~\cite{nadeau_naturebio_2017} and
on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and
energy-harvesting nano-satellites~\cite{kicksat,capybara}.

Energy-harvesting devices create unique challenges because they operate {\em
intermittently} when energy is
available~\cite{hicks_isca_2017,lucia_snapl_2017}. An energy-harvesting device
buffers energy in a small storage capacitor~\cite{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., nW to $\mu$W) compared to a platform's operating
power level (hundreds of $\mu$W to mW). A device operates briefly until it depletes its buffered energy, after which, the device shuts
down and recharges to operate again later---corresponds to {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of
operation-power failure-restart cycles. As an example, recharge times may be
tens of seconds in radio frequency powered medical device~\cite[Fig.
3c]{nadeau_naturebio_2017}. The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor~\cite{capybara} and some devices fail and restart operating $\approx$10 to
$\approx$100 times per second~\cite{tan_infocom_2016,mementos,nvp}.

\textbf{Data Consistency in Intermittent Computing.}  
%Software in an energy-harvesting system operates according to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, which corresponds to the operation-failure-restart cycles of an energy-harvesting device. In an intermittent software execution, an {\em operating period} proceeds for anarbitrary duration (dictated by energy availability) before being interruptedby a {\em failure period}. In a failure period, 
Upon power failures, a device loses the volatile
state in its registers, stack, SRAM, and retains the state of any non-volatile
memory, such as FRAM or Flash. While capturing periodic
checkpoints~\cite{mementos,quickrecall} and sleep
scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution
progress, failures can leave non-volatile state incorrectly, partially updated,
and may leave checkpointed volatile state inconsistent with non-volatile state.
These inconsistencies cause intermittent execution to deviate from
continuously-powered behavior, often leading to an unrecoverable
application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for
intermittently-powered devices: (i) \emph{software-based programming and
execution models}~\cite{dino,ratchet,chain,alpaca} and (ii)
\emph{hardware-based computer architecture
support}~\cite{hicks_isca_2017,idetic,nvp}. Complex architectural changes are
expensive to design, verify, and manufacture. New architectures are also
inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software
approaches are simpler and applicable to existing devices today, but prior
software approaches have a number of limitations that are the focus of the work
in this paper. A key limitation of prior software approaches that we address in this paper is the {\em inflexibility} of a program statically decomposed into tasks.

\begin{figure} %{t!}{0.45\textwidth}
    \centering
    \includegraphics[width=\columnwidth]{figures/intro-figure.pdf}
    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure). On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
    \label{fig:coalesce}
\end{figure}

\textbf{Task Decomposition of Intermittent Programs.} {\em
Task-based} programming and execution models require a
programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to
statically decompose a program into a collection of {\em tasks}.  A task can
include arbitrary computation that should be executed despite arbitrarily-timed power failures.
%and these existing systems guarantee that each
%tasks executes {\em atomically}, despite arbitrarily-timed power failures.  
%
The programmer explicitly expresses task-to-task control flow, or the compiler
converts existing program control-flow into task-to-task control flow to sequence 
the tasks that it defines.
%
Figure~\ref{fig:coalesce} illustrates how a program's tasks execute and
shows how tasks can impose a run-time overhead. At each task transition, the system incurs an overhead to track and atomically \emph{commit} modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}. The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.

@@

%A key challenge is that the length of a software
%task's execution is \emph{limited by the fixed total amount of energy} that a
%device can buffer in hardware. A task's code is static, but the duration of its
%execution may be input-dependent and is difficult to predict. To illustrate
%this, we refer to Figure~\ref{fig:1} showing the execution time of two
%applications (the same application X, but divided into X and Y tasks as
%in~\cite{chain}) running on RFID antenna-powered Computational
%RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna
%distances (near---stable energy supply/far---high energy intermittency). At
%short distance: execution will takes too long caused by runtime cost of
%marshalling excessively small tasks. At far distance: program \emph{might never
%execute} if task execution consumes more energy than the system can buffer.
%This calls for at-runtime adaptive task division of any transiently-powered
%application---namely \emph{task virtualization}.

@@

%variables in non-volatile memory at \emph{run-time} that can
%break the atomicity of the \emph{virtual task}. Consider the
%example depicted in Fig.~\ref{fig:virtualization}: Tasks 1,
%2 and 3 being executed consecutively. All tasks in this
%example are atomic since they do not have WAR dependency on
%the persistent variables they are accessing, e.g. \emph{x}
%is only read and \emph{y} is only written within Task 1.
%Now, suppose three tasks have been virtualized into a single
%one at run-time, namely Task 4: since \emph{x} is now first
%\emph{read} and then \emph{written}, a WAR dependency on
%\emph{x} is introduced dynamically at
%run-time---unfortunately Task 4 is \emph{no more atomic}
%since its re-execution will not always produce the same
%results.
\textbf{Challenges.} While a compiler can use static data privatization and commit instrumentation
(i.e., redo-logging) to eliminate statically identifiable, inter-task data
dependences~\cite{alpaca}, a \emph{dynamic} dependence between two coalesced
tasks requires dynamic privatization and commit actions. 
%
To ensure consistency and respect inter-task dependences, \sys uses a novel
approach \emph{power failure-immune memory virtualization} that buffers non-volatile variable
updates in volatile memory during a coalesced task execution, before committing
them to non-volatile memory at the dynamic task boundary.
%
%Therefore, we require a \emph{new execution
%model} that keeps each virtual task atomic by committing the
%modified persistent variables at the boundary defined
%dynamically at run-time---keeping the non-volatile memory
%unmodified upon a power-interrupt and preserving its
%consistency.
A static task decomposition model assumes that each
single task can execute to completion. If the hardware energy buffer provides
inadequate energy to execute each single task to completion, a program will not
terminate~\cite{cleancut_2018}. To avoid non-termination under adversarial
energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
Partial execution avoids non-termination by committing the intermediate state of a long-running
task that has repeatedly failed and restarted. Partial execution violates
task atomicity, but preserves forward progress; if a programmer knows that
task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
%
% !!!! % It should be emphasized that unlike a voltage-threshold-based protection approaches that may suffer from a non-termination problem if the checkpoint has to be placed within the interrupt-disabled region, a timer-based approach adapts its partial commit location on repeated failures. 
%
%Task coalescing removes the burden from the programmer, because it accepts any
%task decomposition and improves it dynamically. To further reduce the
%programmer effort, we propose a compiler pass for \emph{automatic
%decomposition} of programs into (small) \emph{atomic} tasks. The compiler
%identifies non-volatile variables shared across tasks as tasks are created, and
%instruments reads and writes of those variables, using memory virtualization,
%to keep the data consistent in the presence of power loss.
%
%\textcolor{red}{Despite being limited to a subset of the C language, the
%automatic task decomposition allowed us to port several	applications to an
%intermittent platform with a moderate effort.}
%

### 2018-08-02 paper/asplos19/paper/texFigCode/dma.tex
preparing for asplos19

	\node[punkt, below = -3cm of sram,fill=red!30] (x-sram) {x};
	\node[punkt, below = 0.3cm of x-sram,fill=blue!30] (y-sram) {y};
	\node[punkt, below = 0.3cm of y-sram,fill=blue!30] (z-sram) {z};

@@

	\path[every node/.style={font=\sffamily\small}]
	([yshift=0.3cm] task1.east) edge [->, dotted, thick, bend left] node {(1) get x} (x-sram.west)
	(x-sram.east) edge [->, dotted, thick, bend left] node[xshift=0.5cm] {(2) read x} (x1-fram.west)

@@

	(x-sram.east) edge [->, dotted, thick, bend right] node[xshift=-0.4cm] {(6) commit-I x} (x2-fram.west)
	(z-sram.east) edge [->, dotted, thick, bend right] node[xshift=-0.4cm] {(7) commit-I z} (z2-fram.west)
	(y-sram.east) edge [->, dotted, thick, bend right] node[xshift=-0.4cm] {(8) commit-I y} (y2-fram.west)

	(x2-fram.east) edge [->, dotted, thick, bend right] node[xshift=2.5cm] {(9) commit-II x} (x1-fram.east)
	(y2-fram.east) edge [->, dotted, thick, bend right] node[xshift=2.5cm] {(10) commit-II y} (y1-fram.east)
	(z2-fram.east) edge [->, dotted, thick, bend right] node[xshift=2.5cm] {(11) commit-II z} (z1-fram.east)
	;

>>>>

	\node[punkt, below = -3cm of sram,fill=blue!30] (x-sram) {x};
	\node[punkt, below = 0.3cm of x-sram,fill=blue!30] (y-sram) {y};
	\node[punkt, below = 0.3cm of y-sram,fill=blue!30] (z-sram) {z};

@@

	\path[every node/.style={font=\sffamily\small}]
	 (fram1.west) edge [->, dotted, thick, bend right] node[xshift=-1cm] {(1) flush in (DMA)}([yshift=0.3cm] sram.east)

	([yshift=0.2cm]x-sram.west) edge [->, dotted, thick, bend right] node[xshift=-0.5cm] {(2) read x} ([yshift=0.6cm]task1.east)

@@

	(sram.east) edge [->, dotted, thick, bend right] node[align=center,xshift=-0.5cm] {(6)  flush out \\ (DMA)} (fram2.west)

	(fram2.east) edge [->, dotted, thick, bend right] node[align=center,xshift=1.5cm] {(7) flush out \\ (DMA)} (fram1.east)
	;

### 2018-08-02 paper/asplos19/paper/texFigCode/dynamic-chain-2.tex
preparing for asplos19

	\node[punkt] (task1) {Task (1)};
	\node[punkt, below =0.7cm of task1] (task2) {Task (2)};
	\node[punkt, below =0.7cm of task2] (task3) {Task (3)};
	\node[punkt, below =0.7cm of task3] (task4) {Task (4)};
	\node[punkt, right =0.9cm of task2] (task5) {Task (5)};
	\node[punkt, right =2cm of task4] (task6) {Task (6)};

@@

	\begin{pgfonlayer}{bg}    % select the background layer
		\node[draw,thick,dotted,red,minimum width=7em, fill=blue!10, fit=(task1) (task2) (task3)] (Virtualized) {};
		\draw[<->] ([xshift=-1.7cm]Virtualized.north) --node[sloped, anchor=south] {Virtualization Window} ([xshift=-1.7cm]Virtualized.south);
	\end{pgfonlayer}

>>>>

	\node[punkt] (task1) {Task (1)};
	\node[punkt, below =0.7cm of task1] (task2) {Task (2)};
	\node[punkt, below =0.7cm of task2] (task3) {Task (3)};
	\node[punkt, below =0.7cm of task3] (task4) {Task (4)};
	\node[punkt, right =0.9cm of task2] (task5) {Task (5)};
	\node[punkt, right =1.6cm of task4] (task6) {Task (6)};

@@

	\begin{pgfonlayer}{bg}    % select the background layer
		\node[draw,thick,dotted,red,minimum width=7em, fill=blue!10, fit=(task1) (task2) (task3)] (Virtualized) {};
		\draw[<->] ([xshift=-1.4cm]Virtualized.north) --node[sloped, anchor=south] {} ([xshift=-1.4cm]Virtualized.south);
		% \draw[] ([xshift=-1.5cm]Virtualized.north) --node[sloped, anchor=south] {} ([xshift=-1.5cm]Virtualized.south);
	\end{pgfonlayer}
